Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-08-20 manuscript.Rnw
removed notes on poorter, I think I have accounted for it now

% Notes on Poorter et al. 2015.
% - we have to demonstrate that our work is still novel
% - their focus is firstly on the size-related shifts in biomass partitioning, we don't study that in detail, and can now justify that by referring back to Poorter.
% - they don't have SLA data for the same plants, and in fact predict what we found: 
% "It is also known that, on average, the specific
% leaf area (SLA; leaf area per unit leaf mass) of evergreen woody
% species is 2–3 times higher than that of deciduous species
% (Poorter et al., 2009). Hence, all else being equal, the 2–3 times
% lower SLA in evergreens would have to be compensated by a 2–3
% time higher LMF to arrive at the more or less similar LAI."
% - so basically we can say in the intro that we wish to test this hypothesis (as we already say), and also that Poorter speculated that this may be so (although here they contrast deciduous vs. evergreen, and not angiosperm vs. gymnosperm, so our results are still novel in that case as well)

% - we have to clearly state that our data has *no overlap* whatsoever with Poorter, this is a good thing.
% - their data includes potted plants in glass houses and field-grown, as well as stand biomass data divided by stocking. Our data are more homogenous, with only field-grown plants, individual estimates.
% - we have larger trees included; in fact the largest in the world, we need to state that in the intro to again show the novelty (our max mass is ca. 300t, theirs is 1t)
% - they don't actually show LMF vs. plant size split by PFT (their Fig 2a), so our result is also not scooped there.
% - they found no difference in the LMF residual between evergreen angiosperm and gymnosperm (Fig. 4), or at least I don't think so.

@@

A previous compilation of plant biomass partitioning \citep{poorter2012} (which included many non field-grown plants) did not find a correlation between LMF and LMA across 8 PFTs (including non-woody PFTs). One possible reason for this discrepancy, apart from differences in sample size and growing conditions, is that they did not correct for plant size in their comparison between LMF and LMA. We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale.

% Is it really necessary to invoke differences in LL?

@@

<<results='asis'>>=
print(xtable(table_varpart2, label="table/varpartB",
             caption="Explained variance by plant functional type (PFT), plant height (H) and biome (B), for five whole-plant variables. Each variable was added to a linear mixed-effects model, using species within dataset as a random effect. The $R^2$ shown is that explained by the fixed effects only. All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."),
      include.rownames=FALSE, caption.placement="top", sanitize.text.function=function(x){x})
@

@@

print(xtable(tab, align="lrrrr",
             caption="Explained variance by plant functional type (PFT), plant height (H) and biome (B), for five whole-plant variables. Each variable was added to a linear mixed-effects model, using species within dataset as a random effect. The $R^2$ shown is that explained by the fixed effects only. All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."),
             hline.after=c(0,1,7,8,14),
             include.rownames=FALSE, 
             include.colnames=FALSE, 
             caption.placement="top", sanitize.text.function=function(x){x})
@

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure1.pdf}
    \caption{Overview of the data. (a) Variables were measured on \Sexpr{nrow(dataset)} individual plants from \Sexpr{length(unique(dataset$species))} species. (b) Coverage of the dataset across global climate space. Grey hexagons indicate the number of 0.5\degree~cells with woody vegetation across the space. Colour symbols show the locations of sampled individuals for three dominant woody functional types.}
    \label{fig:figoverview}
\end{figure}

>>>>

@@

We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012,poorter2015}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale, consistent with \citet{poorter2015}.

@@

@@

print(xtable(tab, align="lrrrr",
             caption="Explained variance by plant functional type (PFT), plant height (H) and climate variables (either biome (B), or MAP and MAT) for five whole-plant variables. For the analysis with biome, each variable was added to a linear mixed-effects model, using species within dataset as a random effect. The $R^2$ shown is that explained by the fixed effects only. All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models. For the analysis with continuous climate variables MAP and MAT, each variable was added as a smooth term in a generalized additive model (GAM), with the exception of PFT (a categorical variable). Note that explained variance for the linear mixed-effects model and GAM is not necessarily the same even with the same predictors, due to different methods for fitting and estimating the explained variance."),
             hline.after=c(0,1,7,8,14),
             include.rownames=FALSE, 
             include.colnames=FALSE, 
             caption.placement="top", sanitize.text.function=function(x){x})
@

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure1.pdf}
    \caption{Overview of the data. (a) Variables were measured on up to \Sexpr{nrow(dataset)} individual plants from \Sexpr{length(unique(dataset$species))} species. (b) Coverage of the dataset across global climate space. Grey hexagons indicate the number of 0.5\degree~cells with woody vegetation across the space. Colour symbols show the locations of sampled individuals for three dominant woody functional types.}
    \label{fig:figoverview}
\end{figure}
